---
title: "Lab3"
author: "Soyeon Lee"
date: "3/4/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
#ok to share my example on our course website for this lab

library(tibble)
library(dplyr)

data<-tibble(X=c(4,4,7,7,10,10),
             T=c(1,2,2,4,3,6),
             Y=c(14,23,30,50,39,67))

```

##An alternative view for finding unique variances
# As shown in $The shared part$ section, the total variance predicted by X and T is composed of three sections: the unique part of X, the unique part of T, and the shared variance between X and T. However, the total variance predicted by X and T can also be viewed as the sum of total variance predicted by one variable and the variance unique to the second variable.  

#Unlike the de-correlation method, finding the unique part of each variable does not involve de-correlation.  To obtain the unique part of one variable, the total variance of the second variable is subtracted from the total variance predicted by X and T.    
```{r}
#total variance predicted by X and T
overall_model<-summary(lm(Y~X+T, data=data))$r.squared

#Finding the unique part of X
##total variance of Y predicted by T
lm.yt<-summary(lm(Y~T,data=data))$r.squared

##Extract the unique part of X
uniq_X<-overall_model - lm.yt
uniq_X

##Finding the unique part of T
lm.yx<-summary(lm(Y~X,data=data))$r.squared

##Extract the unique part of T
uniq_T<-overall_model - lm.yx
uniq_T
```
#With a bit of algebra, we can get the variance common to both X and T:
```{r}
X_Tshared<-overall_model - uniq_X - uniq_T
X_Tshared
```
#Since we have the explained portion of the variance of Y, we can also obtain the unexplained variance
```{r}
#explained portion
explained<-1-overall_model
explained
```

