---
title: "Lab4"
author: "Soyeon Lee"
date: "3/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Consider the following example data for a between-subjects experiment with two groups, A and B:

example_data <- tibble(Group = rep(c("A","B"), each = 5),
                       DV = c(2,4,3,5,4,7,6,5,6,7))

Use R to conduct a t.test and ANOVA on this data. Then use R to prove that the results of both analyses are the same. For example, prove that the p-values are the same, and prove that the F-value and T-value are related. (3 points)

```{r}
library(tibble)

example_data <- tibble(Group = rep(c("A","B"), each = 5),
                       DV = c(2,4,3,5,4,7,6,5,6,7))
#t-test result
ttest_result<-t.test(DV~Group, var.equal=TRUE, data=example_data)
ttest_result

#ANOVA result
ANOVA_result<-aov(DV~Group,data=example_data)
my_anova_summary<-summary(ANOVA_result)

#Results of both analyses show that p-values are the same
(ttest_result$p.value)
(my_anova_summary[[1]]$`Pr(>F)`)

ttest_result$p.value==my_anova_summary[[1]]$`Pr(>F)`
```

2. Look at the lab on ANOVA that I wrote for our undergraduate statistics OER lab manual https://crumplab.github.io/statisticsLab/lab-8-one-way-anova.html. That lab shows an example of obtaining data from a published paper in psych science where a one-factor ANOVA was used as a part of the analysis. Load the data, conduct the ANOVA, report a ggplot of the means, and use papaja to help you write a short results section reporting the ANOVA result. (3 points).

```{r}
#load the data
library(data.table)
all_data<-fread("data/Jamesetal2015Experiment2.csv")

#Conduct the ANOVA
##rename the levels with code from statisticsLab/lab-8
all_data$Condition<-as.factor(all_data$Condition)
levels(all_data$Condition)<-c("Control",
                              "Reactivation+Tetris",
                              "Tetris_only",
                              "Reactivation_only")
##rename the long variable "Days_One_to_Seven_Number_of_Intrusions" 
names(all_data)[names(all_data)=="Days_One_to_Seven_Number_of_Intrusions"]<-"Seven_Day_Intrusions"

##use anova function for the analysis 
anova.out<-aov(Seven_Day_Intrusions~Condition,data=all_data)
summary(anova.out)

#ggplot of the means
library(dplyr)
means_by_condition<-all_data%>%
  group_by(Condition)%>%
  summarize(means=mean(Seven_Day_Intrusions))

library(ggplot2)
ggplot(means_by_condition,aes(x=Condition, y=means,fill=Condition))+
  geom_bar(stat="identity")+
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

#use papaja to report the result
library(papaja)
apa_print(anova.out)

```

