---
title: "Lab4"
author: "Soyeon Lee"
date: "3/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Consider the following example data for a between-subjects experiment with two groups, A and B:

example_data <- tibble(Group = rep(c("A","B"), each = 5),
                       DV = c(2,4,3,5,4,7,6,5,6,7))

Use R to conduct a t.test and ANOVA on this data. Then use R to prove that the results of both analyses are the same. For example, prove that the p-values are the same, and prove that the F-value and T-value are related. (3 points)
#Independence:90

```{r}
library(tibble)

example_data <- tibble(Group = rep(c("A","B"), each = 5),
                       DV = c(2,4,3,5,4,7,6,5,6,7))
#t-test 
my_t<-t.test(DV~Group, var.equal=TRUE, data=example_data)
my_t

#ANOVA 
my_anova_summary<-summary(aov(DV~Group,data=example_data))

#Results of both analyses show that p-values are the same
##anova_p_value is a vector with two values 
ttest_p_value<-my_t$p.value
anova_p_value<-my_anova_summary[[1]]$`Pr(>F)`[1]

round(ttest_p_value,5) == round(anova_p_value,5)

#Show that F and t statistics are related
t_squared<-my_t$statistic^2
F_value<-my_anova_summary[[1]]$`F value` [1]

round(t_squared,1)==round(F_value,1)
```

2. Look at the lab on ANOVA that I wrote for our undergraduate statistics OER lab manual https://crumplab.github.io/statisticsLab/lab-8-one-way-anova.html. That lab shows an example of obtaining data from a published paper in psych science where a one-factor ANOVA was used as a part of the analysis. Load the data, conduct the ANOVA, report a ggplot of the means, and use papaja to help you write a short results section reporting the ANOVA result. (3 points).
Independence: 80

```{r}
#load the data
library(data.table)
all_data<-fread("data/Jamesetal2015Experiment2.csv")

#Conduct the ANOVA
##rename the levels with code from statisticsLab/lab-8
all_data$Condition<-as.factor(all_data$Condition)
levels(all_data$Condition)<-c("Control",
                              "Reactivation+Tetris",
                              "Tetris_only",
                              "Reactivation_only")
##Shorten the DV name "Days_One_to_Seven_Number_of_Intrusions" 
names(all_data)[names(all_data)=="Days_One_to_Seven_Number_of_Intrusions"]<-"Seven_Day_Intrusions"

#ggplot of the means
##means calculated using the fun="mean"
library(ggplot2)
ggplot(all_data,aes(x=Condition,y=Seven_Day_Intrusions))+
  geom_bar(stat="summary",
           fun="mean",
           position="dodge",
           aes(fill=Condition))+
  geom_point()+
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

##use anova function for the analysis 
anova.out<-aov(Seven_Day_Intrusions~Condition,data=all_data)
summary(anova.out)
  
#use papaja to report the result
library(papaja)
apa_print(anova.out)
```

#APA style report
The one-factor between-subjects ANOVA revealed that intervention type (No-task control, Reactivation Plus tetris, Tetris only, Reactivation only) influences the number of intrusive memories, `r apa_print(anova.out)$full_result$Condition`.
