---
title: "Lab8"
author: "Soyeon Lee"
date: "4/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Consider a 2x2 design. Assume the DV is measured from a normal distribution with mean 0, and standard deviation 1. Assume that the main effect of A causes a total shift of .5 standard deviations of the mean between the levels. Assume that level 1 of B is a control, where you expect to measure the standard effect of A. Assume that level 2 of B is an experimental factor intended to reduce the effect of A by .25 standard deviations.

A. create a ggplot2 figure that depicts the expected results from this design (2 points)
Independence: 50
```{r}
#load libraries
library(tibble)
library(ggplot2)

# define 2-factor score model
grand_mean <- 0
A <- c(0,.5)
B <- c(0,0)
AB <- c(0,0,0,-.25)

# create design table
model_data <- tibble()
for(i in 1:length(A)){
  for(j in 1:length(B)){
    IVA <- i
    IVB <- j
    DV <- grand_mean+A[i]+B[j]+AB[(i-1)*length(B)+j]
    sc_GM <- grand_mean
    sc_A <- A[i]
    sc_B <- B[j]
    sc_AB <- AB[(i-1)*length(B)+j]
    row_entry <- tibble(IVA,IVB,DV,
                        sc_GM,sc_A,sc_B,sc_AB)
    model_data <- rbind(model_data,row_entry)
  }
}

knitr::kable(model_data)

# generate a bar graph
ggplot(model_data, aes(y=DV,
                       x=as.factor(IVA),
                       fill=as.factor(IVB)))+
       geom_bar(stat='identity', position='dodge')
```


```{r}
##Alternative solution for Question A

model_data<-tibble(IVA=c(1,1,2,2),
                   IVB=c(1,2,1,2),
                   DV=c(0,0,.5,.25))

ggplot(model_data, aes(y=DV,
                       x=IVA,
                       linetype=as.factor(IVB)))+
       geom_line()+
       geom_point()
```

Conduct a simulation-based power analysis to answer the questions.

B. How many subjects are needed to detect the main effect of A with power = .8? (2 points)
Independence: 80
```{r}
# N per condition selected by trial and error 
N <- 58 
A_pvalue <- c()

for(i in 1:1000){
  IVA <- rep(rep(c("1","2"), each=2),N)
  IVB <- rep(rep(c("1","2"), 2),N)
  DV <- c(replicate(N,c(rnorm(1,0,1), # means A1B1
                        rnorm(1,0,1), # means A1B2
                        rnorm(1,.5,1), # means A2B1
                        rnorm(1,.25,1)  # means A2B2
          )))
  sim_df <- data.frame(IVA,IVB,DV)

  aov_results <- summary(aov(DV~IVA*IVB, sim_df))
  A_pvalue[i]<-aov_results[[1]]$`Pr(>F)`[1]
}

length(A_pvalue[A_pvalue<0.05])/1000
```


```{r}
##Alternative solution for Question B
N <- 58 
A_pvalue <- c()

for(i in 1:1000){
  sim_df<-tibble(
    IVA = rep(rep(c("1","2"), each=2),N),
    IVB = rep(rep(c("1","2"), 2),N),
    DV = rnorm(4*N, c(0,0,.5,.25),1)
  )
  aov_results <- summary(aov(DV~IVA*IVB, sim_df))
  
  A_pvalue[i]<-aov_results[[1]]$`Pr(>F)`[1]
}

length(A_pvalue[A_pvalue<0.05])/1000
```

C. How many subjects are needed to detect the interaction effect with power = .8? (2 points)
Independence: 90
```{r}
#N per condition selected by trial and error
N <- 500 
AB_pvalue <- c()

for(i in 1:1000){
  sim_df<-tibble(
    IVA = rep(rep(c("1","2"), each=2),N),
    IVB = rep(rep(c("1","2"), 2),N),
    DV = rnorm(4*N, c(0,0,.5,.25),1)
  )
  aov_results <- summary(aov(DV~IVA*IVB, sim_df))
  
  AB_pvalue[i]<-aov_results[[1]]$`Pr(>F)`[3]
}

length(AB_pvalue[AB_pvalue<0.05])/1000    
```


Bonus point question:

B1. Create a power curve showing how power for the interaction effect in this example is influenced by number of subjects. Choose a range of N from 25 to 800 (per cell) and run a simulation-based power analysis for increments of 25 subjects. Then plot the results using ggplot2 (2 points).
Independence:90
```{r}



N <- 500 
AB_pvalue <- c()

for(i in 1:1000){
  sim_df<-tibble(
    IVA = rep(rep(c("1","2"), each=2),N),
    IVB = rep(rep(c("1","2"), 2),N),
    DV = rnorm(4*N, c(0,0,.5,.25),1)
  )
  aov_results <- summary(aov(DV~IVA*IVB, sim_df))
  
  AB_pvalue[i]<-aov_results[[1]]$`Pr(>F)`[3]
}

length(AB_pvalue[AB_pvalue<0.05])/1000    
```

