---
title: "Lab9"
author: "Soyeon Lee"
date: "5/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Create an R script that can generate simulated data for the following repeated measures design. (2 points)

A. The dependent variable is assumed to come from a normal distribution with mean = 0 and standard deviation = 1.

B. There is one repeated measures factor with 5 levels (Down1, Down2, Control, Up1, Up2). The control group is assumed to have no effect. The Down1 and Down2 levels shift the mean down by 1 and 2 standard deviations, respectively. The Up1 and Up2 levels shift the mean up by 1 and 2 standard deviations, respectively.

C. There are 6 subjects in the experiment, and they are each measured once in each condition. The 6 subjects are assumed to be different from one another (e.g., they will have different baseline means in the control condition), but they will all be influenced by the IV in the exact same way (e.g., no interaction).

Independence: 10
```{r}
library(tibble)
library(dplyr)
library(ggplot2)

sim_data<-tibble(
  subjects=rep(1:6,each=5),
  IV=rep(c("Down2","Down1","Control","Up1","Up2"),times=6),
  DV=rnorm(5*6,c(-2,-1,0,1,2),1)
  )%>%
    mutate(DV = DV+rep(rnorm(6,0,1),each=5))

sim_data$IV<-factor(sim_data$IV,
                    levels=c("Down2","Down1","Control","Up1","Up2"))

sim_data$subjects<-as.factor(sim_data$subjects)

ggplot(sim_data, aes(x=IV, y=DV, group=subjects, color=subjects))+
  geom_point()+
  geom_line()
                 
```

2. Run a simulation to determine the proportion of experiments that would return a significant result for the above design. Assume that the effect of the levels of the IV are increments of .1 of a standard deviation, rather than increments of 1 as in the above design. (2 points?)
Independence: 70
```{r}
p_value<-c()

for (i in 1:1000){
  
  sim_data<-tibble(
    subjects=rep(1:6,each=5),
    IV=rep(c("Down2","Down1","Control","Up1","Up2"),times=6),
    DV=rnorm(5*6,c(-.2,-.1,0,.1,.2),1)
  )%>%
    mutate(DV = DV+rep(rnorm(6,0,1),each=5))

  sim_data$IV<-factor(sim_data$IV,
                      levels=c("Down2","Down1","Control","Up1","Up2"))

  sim_data$subjects<-as.factor(sim_data$subjects)
  
  aov_out<-summary(aov(DV~IV + Error(subjects),sim_data))

  p_value[i]<-aov_out[2]$`Error: Within`[[1]]$`Pr(>F)`[1] 
}

#get proportion of p values less than .05
length(p_value[p_value<.05])/1000
```

3. Demonstrate that the Godden and Baddeley example data from the textbook (19.5), which used a 2x2 repeated measures design, can be be analyzed with one-sample t-tests to return the same results. Specifically, show the one-sample t-tests for each main effect and the interaction. (2 points)
Independence: 10
```{r}
godden_baddeley <- tribble(~Subjects,~LearningPlace,~TestingPlace,~Recall,
        "s1","On Land","On Land",34,
        "s2","On Land","On Land",37,
        "s3","On Land","On Land",27,
        "s4","On Land","On Land",43,
        "s5","On Land","On Land",44,
        "s1","On Land","Under Sea",18,
        "s2","On Land","Under Sea",21,
        "s3","On Land","Under Sea",25,
        "s4","On Land","Under Sea",37,
        "s5","On Land","Under Sea",34,
        "s1","Under Sea","On Land",14,
        "s2","Under Sea","On Land",21,
        "s3","Under Sea","On Land",31,
        "s4","Under Sea","On Land",27,
        "s5","Under Sea","On Land",32,
        "s1","Under Sea","Under Sea",22,
        "s2","Under Sea","Under Sea",25,
        "s3","Under Sea","Under Sea",33,
        "s4","Under Sea","Under Sea",33,
        "s5","Under Sea","Under Sea",42
        )

# convert IVs to factors
godden_baddeley <- godden_baddeley %>%
  mutate(Subjects = as.factor(Subjects),
         LearningPlace = as.factor(LearningPlace),
         TestingPlace = as.factor(TestingPlace))


####one-sample t test solutions

##Main effect of Learning Place
#sort out the Learning Place means
learning_place_means<-godden_baddeley%>%
    group_by(Subjects,LearningPlace)%>%
    summarize(mean_recall=mean(Recall))

#paired-sample t test 
t.test(mean_recall~LearningPlace,
       paired=TRUE,
       data=learning_place_means)


##Main effect of Testing Place 
#sort out the main effect of Testing Place
testing_place_means<-godden_baddeley%>%
  group_by(Subjects,TestingPlace)%>%
  summarize(mean_recall=mean(Recall))

#sort out the Testing Place means
Testing_land<-testing_place_means%>%
  filter(TestingPlace=="On Land")%>%
  select(mean_recall)

Testing_sea<-testing_place_means%>%
  filter(TestingPlace=="Under Sea")%>%
  select(mean_recall)

#one sample t test 
t.test(Testing_land$mean_recall -
       Testing_sea$mean_recall,
       mu=0)


##Interaction solution
#the effect of testing On Land for learning On Land
LL <- godden_baddeley%>%
        filter(LearningPlace=="On Land",
               TestingPlace=="On Land")%>%
        pull(Recall)

LS<-godden_baddeley%>%
        filter(LearningPlace=="On Land",
               TestingPlace=="Under Sea")%>%
        pull(Recall)


#the effect of testing On Land for learning Under Sea
SL<-godden_baddeley%>%
        filter(LearningPlace=="Under Sea",
               TestingPlace=="On Land")%>%
        pull(Recall)

SS <- godden_baddeley%>%
        filter(LearningPlace=="Under Sea",
               TestingPlace=="Under Sea")%>%
        pull(Recall)


#single sample t test 
t.test((LL - LS)-(SL - SS), mu=0)
```

Bonus Points

These bonus questions are about the sphericity concept, and they will involve modifying the example data from the textbook (same as used in concept sectdion 1). The data is reprinted here for your convenience:
```{r}
sphericity <- tribble(~S, ~a1, ~a2, ~a3, ~a4,
        "s1",76,64,34,26,
        "s2",60,48,46,30,
        "s3",58,34,32,28,
        "s4",46,46,32,28,
        "s5",30,18,36,28
        )
```

4. Create a line plot showing how each of the 5 subjects perform across the levels of the IV. Discuss how the line plot visually shows the sphericity problem in the data (1 point)
Independence: 80
```{r}
library(tidyr)
library(dplyr)

#convert it into a long data and plot
sphericity<-pivot_longer(sphericity,
                         cols=!S,
                         names_to="IV",
                         values_to="DV")

sphericity<-sphericity%>%
  mutate(S=as.factor(S),
         IV=as.factor(IV))

ggplot(sphericity, aes(x=IV, 
                       y=DV, 
                       color=S,
                       group=S))+
  geom_point()+
  geom_line()

##Explanation
#If there is sphericity, a pattern across subjects on one level should be parallel to show that there is same effect at each level.  Here, we do not see that there is sphericity as the lines representing the results of each subject are very different--lines between levels converge or cross. 
```


5. Modify the above data to remove the sphericity problem. Specifically, ensure that all subjects are different from another (there overall means are different), and that the IV has the same effect for each level and each subject (no interaction). Then, plot the new data and discuss how the graph shows the sphericity problem has been removed. (1 point)
Independence:10
```{r}
#correct values to have same effect at each level for all subjects
sphericity <- tribble(~S, ~a1, ~a2, ~a3, ~a4,
        "s1",76,64,34,26,
        "s2",60,48,18,10,
        "s3",58,46,16,8,
        "s4",46,34,4,-4,
        "s5",30,18,-12,-20
        )

sphericity<-pivot_longer(sphericity,
                         cols=!S,
                         names_to="IV",
                         values_to="DV")

ggplot(sphericity, aes(x=IV, 
                       y=DV, 
                       color=S,
                       group=S))+
  geom_point()+
  geom_line()

```

6. Calculate the Greenhouse-Geisser estimate of epsilon for your modified data to demonstrate you have removed the sphericity problem. (1 point)
Independence:80
```{r}
#sphericity corrected here
sphericity <- tribble(~S, ~a1, ~a2, ~a3, ~a4,
        "s1",76,64,34,26,
        "s2",60,48,18,10,
        "s3",58,46,16,8,
        "s4",46,34,4,-4,
        "s5",30,18,-12,-20
        )
textbook<-sphericity

#double-centering
cov_matrix <- cov(textbook[,2:5])
col_mean_matrix <- cov_matrix*0 + colMeans(cov_matrix)
row_mean_matrix <- t(cov_matrix*0 + rowMeans(cov_matrix))
dc_matrix <- cov_matrix - col_mean_matrix -row_mean_matrix + mean(cov_matrix)

## greenhouse-geisser
sum(diag(dc_matrix))^2 / ((dim(dc_matrix)[1]-1)*sum(dc_matrix^2))

```

